Lessons learned: creating tutorials teaching agent-based modelling for archaeologists

### 2025-03-17 paper.Rmd
update paper based on reviews

---
title: "Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists"
authors:
  - name: Ronald M. Visser
    orcid_id: 0000-0001-6966-1729
    roles: [writing, editing, analysis, visualization]
    affiliation: Saxion University of Applied Sciences
  - name: Karsten Lambers 
    roles: [writing, editing]
    orcid_id: 0000-0001-6432-0925
    affiliation: Leiden University
  - name: Kenneth Aitchison 
    roles: [writing, editing, software, project administration]
    orcid_id: 0000-0001-6594-1408
    affiliation: Landward Research Teoranta
  - name: Annemarie Jutte
    orcid_id: 0009-0004-9322-0674
    roles: [writing, editing, software]
    affiliation: Saxion University of Applied Sciences
  - name: Doug Rocks-Macqueen 
    orcid_id: 0000-0001-9021-0260
    roles: [writing, editing, software]
    affiliation: Landward Research Teoranta
  - name: Laura van der Knaap 
    roles: [writing, editing, software]
    orcid_id: 0000
    affiliation: Leiden University
  - name: Iza Romanowska 
    orcid_id: 0000-0002-9487-2111
    roles: [writing, editing]
    affiliation: Aarhus University
  - name: Tom Brughmans
    roles: [writing, editing]
    orcid_id: 0000-0002-1589-7768
    affiliation: Aarhus University
  - name: Jeroen Linssen
    orcid_id: 0000-0002-2626-1837
    roles: [writing, editing]
    affiliation: Saxion University of Applied Sciences
  - name: Adam Pažout
    roles: [writing, editing]
    orcid_id: 0000-0001-7745-5634
    affiliation: Aarhus University
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  distill::distill_article:
    number_sections: yes
    fig_caption: yes
  bookdown::pdf_document2:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
  bookdown::word_document2:
    number_sections: yes
    fig_caption: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

@@

While Agent-Based Modelling importance in archaeological practice is steadily growing [@romanowska2023] there are significant barriers to its wider adoption. One of the problems is the access to vocational training required to gain the necessary computer programming and modelling skills. A survey of practitioners [@davies2018] found significant deficiencies in the provision of ABM training; 76% of modellers were primarily self-taught with some degree of secondary support, i.e. 27% relied on peer support, only 7% had any sort of formal training at a workshop or summer school, while only 14% received training as part of a degree.

@@

@@

@@

# Method

The project team developed five online tutorials. The tutorials were first designed using storyboards and later converted into interactive guided walk-throughs using NetLOGO Web (<https://www.netlogoweb.org/launch>) and Javascript and hosted on GitHub (<https://github.com/ABMArchaeologists/ABMA_tutorials>). During the project, initial testing was done by all authors and the issues functionality in GitHub was used to tackle bugs.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team (Table \@ref(tab:events)). Recognising the difficulty of delivering practical training in an online environment (Microsoft Teams and Zoom) during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics (<https://www.qualtrics.com/>). The participants of the events were asked to respond to two surveys: one before the start of the event and one after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was shared at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

@@

@@

## Workhops and tutorials

## Final tutorials

@@

## Before the workshops

@@

## During the workshop

@@

## After the workshop

@@

The respondents generally liked the interactive step-by-step way of going through the tutorials, but also valued the interaction with the teachers strongly. The easy and intuitive introduction in ABM and NetLOGO were also mentioned quite often. This was also observed during the workshops as most participants were working at their own pace and did not need much help. From a didactic point of view, the self-guided and self-paced way of going through the tutorials is very efficient for teachers and limits the frustration of students that may otherwise go too fast or fall behind the class when the tempo is dictated by the teacher.

@@

@@

Besides, the tutorial, the website provides: a link that can be used for feedback, basic explanations of agent-based modelling, an introductory presentation from one of the workshops, videos from an online course at Leiden University [@scherjon2019] and external resources that people can use to learn agent-based modelling. The website was mainly developed by a student group from Saxion University of Applied Sciences, see the acknowledgements.

>>>>

---
title: "Lessons learned: creating tutorials teaching agent-based modelling for archaeologists"
authors:
  - name: Ronald M. Visser
    orcid_id: 0000-0001-6966-1729
    roles: [writing, editing, analysis, visualization]
    affiliation: Saxion University of Applied Sciences
  - name: Karsten Lambers 
    roles: [writing, editing]
    orcid_id: 0000-0001-6432-0925
    affiliation: Leiden University
  - name: Kenneth Aitchison 
    roles: [writing, editing, software, project administration]
    orcid_id: 0000-0001-6594-1408
    affiliation: Landward Research Teoranta
  - name: Annemarie Jutte
    orcid_id: 0009-0004-9322-0674
    roles: [writing, editing, software]
    affiliation: Saxion University of Applied Sciences
  - name: Doug Rocks-Macqueen 
    orcid_id: 0000-0001-9021-0260
    roles: [writing, editing, software]
    affiliation: Landward Research Teoranta
  - name: Laura van der Knaap 
    roles: [writing, editing, software]
    orcid_id: 0000
    affiliation: Leiden University
  - name: Iza Romanowska 
    orcid_id: 0000-0002-9487-2111
    roles: [writing, editing]
    affiliation: Aarhus University
  - name: Tom Brughmans
    roles: [writing, editing]
    orcid_id: 0000-0002-1589-7768
    affiliation: Aarhus University
  - name: Jeroen Linssen
    orcid_id: 0000-0002-2626-1837
    roles: [writing, editing]
    affiliation: Saxion University of Applied Sciences
  - name: Adam Pažout
    roles: [writing, editing]
    orcid_id: 0000-0001-7745-5634
    affiliation: Aarhus University
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  distill::distill_article:
    number_sections: yes
    fig_caption: yes
  bookdown::pdf_document2:
    number_sections: yes
    keep_tex: true
    fig_caption: yes
  bookdown::word_document2:
    number_sections: yes
    fig_caption: yes
bibliography: references.bib
csl: journal-of-computer-applications-in-archaeology.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

@@

While agent-based modelling importance in archaeological practice is steadily growing [@romanowska2023] there are significant barriers to its wider adoption. One of the problems is the access to vocational training required to gain the necessary computer programming and modelling skills [@daems2020]. A survey of practitioners [@davies2018] found significant deficiencies in the provision of ABM training; 76% of modellers were primarily self-taught with some degree of secondary support, i.e. 27% relied on peer support, only 7% had any sort of formal training at a workshop or summer school, while only 14% received training as part of a degree.

@@

# Methods

## Open educational resources

@@

## Objectives and target audience

@@

## Online tutorials

The project team developed five online tutorials. The tutorials were first designed using storyboards and later converted into interactive guided walk-throughs using NetLogo Web (<https://www.netlogoweb.org/launch>) and Javascript and hosted on GitHub (<https://github.com/ABMArchaeologists/ABMA_tutorials>). During the project, initial testing was done by all authors and the issues functionality in GitHub was used to tackle bugs.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team (Table \@ref(tab:events)). Recognising the difficulty of delivering practical training in an online environment (Microsoft Teams and Zoom) during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics (<https://www.qualtrics.com/>). The participants of the events were asked to respond to one survey before the start of the event and one after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was shared at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

@@

Besides, the tutorial, the website provides: a link that can be used for feedback, basic explanations of agent-based modelling, an introductory presentation from one of the workshops, videos from an online course at Leiden University [@scherjon2019] and external resources that people can use to learn agent-based modelling. The website was mainly developed by a student group from Saxion University of Applied Sciences [@jutte2024].

@@

## Tutorials

@@

## Tutorials and workshops

The tutorials were tested during various workshops and updated afterward. As mentioned earlier, the participants of these workshops were asked to fill in a questionnaire both before and after the workshop.

### Before the workshops

@@

### During the workshop

@@

### After the workshop

@@

The respondents generally liked the interactive step-by-step way of going through the tutorials, but also valued the interaction with the teachers strongly. The easy and intuitive introduction in ABM and NetLogo were also mentioned quite often. This was also observed during the workshops as most participants were working at their own pace and did not need much help. From a didactic point of view, the self-guided and self-paced way of going through the tutorials is very efficient for teachers and limits the frustration of students that may otherwise go too fast or fall behind the class when the tempo is dictated by the teacher.

@@

# Discussion

@@

## Iterative creation of tutorials

*From review: Although the point can be somewhat interpreted throughout the manuscript, I would strongly suggest that the authors make a brief discussion of the experience (i.e. the iterative design and testing of educational material),*

## Computational skills in archaeology

*From review: particularly in light of shortcomings and resistances to digital (and computational) skills in archaeology.*

### 2025-03-17 paper.tex
update paper based on reviews

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists}
\author{}
\date{\vspace{-2.5em}November 13, 2024}

@@

While Agent-Based Modelling importance in archaeological practice is steadily growing (Romanowska and Scherjon 2023) there are significant barriers to its wider adoption. One of the problems is the access to vocational training required to gain the necessary computer programming and modelling skills. A survey of practitioners (Davies and Romanowska 2018) found significant deficiencies in the provision of ABM training; 76\% of modellers were primarily self-taught with some degree of secondary support, i.e.~27\% relied on peer support, only 7\% had any sort of formal training at a workshop or summer school, while only 14\% received training as part of a degree.

@@

@@

@@

\hypertarget{method}{%
\section{Method}\label{method}}

The project team developed five online tutorials. The tutorials were first designed using storyboards and later converted into interactive guided walk-throughs using NetLOGO Web (\url{https://www.netlogoweb.org/launch}) and Javascript and hosted on GitHub (\url{https://github.com/ABMArchaeologists/ABMA_tutorials}). During the project, initial testing was done by all authors and the issues functionality in GitHub was used to tackle bugs.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team (Table \ref{tab:events}). Recognising the difficulty of delivering practical training in an online environment (Microsoft Teams and Zoom) during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics (\url{https://www.qualtrics.com/}). The participants of the events were asked to respond to two surveys: one before the start of the event and one after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was shared at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

@@

@@

\hypertarget{workhops-and-tutorials}{%
\subsection{Workhops and tutorials}\label{workhops-and-tutorials}}

\hypertarget{final-tutorials}{%
\subsection{Final tutorials}\label{final-tutorials}}

@@

\hypertarget{before-the-workshops}{%
\subsection{Before the workshops}\label{before-the-workshops}}

@@

\hypertarget{during-the-workshop}{%
\subsection{During the workshop}\label{during-the-workshop}}

@@

\hypertarget{after-the-workshop}{%
\subsection{After the workshop}\label{after-the-workshop}}

@@

@@

Besides, the tutorial, the website provides: a link that can be used for feedback, basic explanations of agent-based modelling, an introductory presentation from one of the workshops, videos from an online course at Leiden University (Scherjon, Romanowska and Lambers 2019) and external resources that people can use to learn agent-based modelling. The website was mainly developed by a student group from Saxion University of Applied Sciences, see the acknowledgements.

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Lessons learned: creating tutorials teaching agent-based modelling for archaeologists},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Lessons learned: creating tutorials teaching agent-based modelling for archaeologists}
\author{}
\date{\vspace{-2.5em}March 17, 2025}

@@

While agent-based modelling importance in archaeological practice is steadily growing (Romanowska and Scherjon 2023) there are significant barriers to its wider adoption. One of the problems is the access to vocational training required to gain the necessary computer programming and modelling skills (Daems 2020). A survey of practitioners (Davies and Romanowska 2018) found significant deficiencies in the provision of ABM training; 76\% of modellers were primarily self-taught with some degree of secondary support, i.e.~27\% relied on peer support, only 7\% had any sort of formal training at a workshop or summer school, while only 14\% received training as part of a degree.

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{open-educational-resources}{%
\subsection{Open educational resources}\label{open-educational-resources}}

@@

\hypertarget{objectives-and-target-audience}{%
\subsection{Objectives and target audience}\label{objectives-and-target-audience}}

@@

\hypertarget{online-tutorials}{%
\subsection{Online tutorials}\label{online-tutorials}}

The project team developed five online tutorials. The tutorials were first designed using storyboards and later converted into interactive guided walk-throughs using NetLogo Web (\url{https://www.netlogoweb.org/launch}) and Javascript and hosted on GitHub (\url{https://github.com/ABMArchaeologists/ABMA_tutorials}). During the project, initial testing was done by all authors and the issues functionality in GitHub was used to tackle bugs.

@@

After initial debugging by the project team, the tutorials were tested on a wider sample of participants during several online and in-person events all of which were run by members of the project team (Table \ref{tab:events}). Recognising the difficulty of delivering practical training in an online environment (Microsoft Teams and Zoom) during the online events several members of the team were present which enabled participants to go to break-out rooms with an expert if they needed additional help or wanted to discuss things. Throughout the project we sought feedback from the participants and used it to improve the tutorials prior to the following events. The feedback was obtained through two surveys using Qualtrics (\url{https://www.qualtrics.com/}). The participants of the events were asked to respond to one survey before the start of the event and one after the end of the event. The first survey focused on the background of the participants and their original level of knowledge in relation to ABM (see appendix 1 for the questions). At the end of the event the participants were asked to answer a second survey. Questions of this survey were aimed at measuring the effectiveness of the tutorials and getting feedback on the workshop and tutorials (see appendix 2 for the questions). For some events participants had to register beforehand and we were able to send the pre-workshop survey by email. At other events no registration was possible or necessary and the pre-workshop survey was shared at the start of the event. The post-workshop survey was distributed using QR-codes or links at the end of the event.

@@

Besides, the tutorial, the website provides: a link that can be used for feedback, basic explanations of agent-based modelling, an introductory presentation from one of the workshops, videos from an online course at Leiden University (Scherjon, Romanowska and Lambers 2019) and external resources that people can use to learn agent-based modelling. The website was mainly developed by a student group from Saxion University of Applied Sciences (Jutte et al. 2024).

@@

\hypertarget{tutorials}{%
\subsection{Tutorials}\label{tutorials}}

@@

\hypertarget{tutorials-and-workshops}{%
\subsection{Tutorials and workshops}\label{tutorials-and-workshops}}

The tutorials were tested during various workshops and updated afterward. As mentioned earlier, the participants of these workshops were asked to fill in a questionnaire both before and after the workshop.

\hypertarget{before-the-workshops}{%
\subsubsection{Before the workshops}\label{before-the-workshops}}

@@

\hypertarget{during-the-workshop}{%
\subsubsection{During the workshop}\label{during-the-workshop}}

@@

\hypertarget{after-the-workshop}{%
\subsubsection{After the workshop}\label{after-the-workshop}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{iterative-creation-of-tutorials}{%
\subsection{Iterative creation of tutorials}\label{iterative-creation-of-tutorials}}

\emph{From review: Although the point can be somewhat interpreted throughout the manuscript, I would strongly suggest that the authors make a brief discussion of the experience (i.e.~the iterative design and testing of educational material),}

\hypertarget{computational-skills-in-archaeology}{%
\subsection{Computational skills in archaeology}\label{computational-skills-in-archaeology}}

\emph{From review: particularly in light of shortcomings and resistances to digital (and computational) skills in archaeology.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-daems2020}{}}%
Daems, D. 2020 A Review and Roadmap of Online Learning Platforms and Tutorials in Digital Archaeology. \emph{Advances in Archaeological Practice} 8(1): 87--92. DOI: https://doi.org/\href{https://doi.org/10.1017/aap.2019.47}{10.1017/aap.2019.47}.
